What We Eval in the Shadows

### 2021-08-12 paper/main.tex
Remove expression type explanations from Usage of Eval
Update the comparison to Morandat et al

\citet{ecoop12} had a few paragraphs on the usage of \eval in R. They found the
that \eval is widely used in R code with 8500 call sites in 1238 CRAN packages
and 2.2 million dynamic calls. The 15 most frequent call sites account for 88\%
of those. The \k{match.arg} function is the highest used one with 54\% of all
calls. In the other call sites, they saw two use cases. The most common is the
evaluation of the source code of a promise retrieved by \k{substitute} in a new
environment, \eg as done in the \k{with} function. The other use case is the
invocation of a function whose name or arguments are determined dynamically. For
this purpose, R provides \k{do.call} and thus \eval is overkill. Our corpus is
larger, with \CranPackages packages, \packageTriggeredpkgs packages with at
least one \eval, for a total of \PkgHitEvalCallSites sites and
\packageAllcallsRnd dynamic calls, although they also looked at the Bioconductor
package repository in addition to CRAN. We also track much more runtime
information on the \eval calls during tracing. It enables us to have a more
nuanced and accurate picture, and identify more diverse use cases of \eval.

>>>>

\citet{ecoop12} in their evaluation of the design of the R language, briefly
looked at the use of \eval in their corpus (1238 CRAN packages, R base
libraries, and 515 Bioconductor packages). They found 8500 \eval call sites and
recorded 2.2M \eval calls at runtime. Because they included base, the results
were dominated by a few functions (15 call sites accounted for 88\% of \eval
calls) with \c{match.arg}\footnote{A popular base function that matches an
  argument against a set of candidate values that uses \eval to get the
candidate values.} being responsible for over half of the recorded \eval
calls. They discussed two use cases. The first one is the evaluation of an
expression extracted from a promise using the \k{substitute} call in a new
environment. The other is the invocation of a function with dynamically
computed names and arguments.

In this work, we significantly expand that study. First, our corpus covers
\CranPackages packages (about the whole of CRAN at time of writing) with
\PkgHitEvalCallSites \eval call sites and \packageAllcallsRnd \eval calls.
Next, we track the number of information such as the shapes of expressions
passed to \eval, their provenance, complexity, and side effects. This enables
us to have an accurate and detailed picture of the use of \eval in the R
ecosystem. We use this to provide both quantitative and qualitative views on
\eval, including several use cases to demonstrate the popularity, diversity and
powerfulness of \eval in R.
